---
title: "FDI Data Checks"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(ggplot2)
library(knitr)
library(rworldmap)
library(rnaturalearth)
library(rnaturalearthdata)

source("Functions.R")

```

## Introduction

This project carries out some checks on data that is formatted for the FDI data call  - it does not perform the validation checks that are already performed by the FDI Data Upload portal but instead tries to look at the data and present plots etc to help the user decide whether the information content is correct.

## Load the data

First we'll load the data from the Excel files using the "readxl" package.  Your data files need to already be saved in a sub-directory of this project called "data".  We'll read in all columns as text to avoid problems when there is a NK in a numeric column.

```{r loaddata}

# Don't automatically convert strings to factors
options(stringsAsFactors = FALSE)

# Uncomment the next statement to remove everything from the workspace and start with a clean environment
#rm(list = ls())


myTables <- list(Table_A = read_excel("data/TABLE_A_CATCH.xlsx", col_types = "text")
                 ,Table_B  =  read_excel("data/TABLE_B_REFUSAL_RATE.xlsx", col_types = "text")
                 ,Table_C_NAO  =  read_excel("data/TABLE_C_NAO_OFR_DISCARDS_AGE.xlsx", col_types = "text")
                 ,Table_D_NAO  =  read_excel("data/TABLE_D_NAO_OFR_DISCARDS_LENGTH.xlsx", col_types = "text")
                 ,Table_E_NAO  =  read_excel("data/TABLE_E_NAO_OFR_LANDINGS_AGE.xlsx", col_types = "text")
                 ,Table_F_NAO  =  read_excel("data/TABLE_F_NAO_OFR_LANDINGS_LENGTH.xlsx", col_types = "text")
                 ,Table_G  =  read_excel("data/TABLE_G_EFFORT.xlsx", col_types = "text")
                 ,Table_H  =  read_excel("data/TABLE_H_LANDINGS_BY_RECTANGLE.xlsx", col_types = "text")
                 ,Table_I  =  read_excel("data/TABLE_I_EFFORT_BY_RECTANGLE.xlsx", col_types = "text")
                 ,Table_J  =  read_excel("data/TABLE_J_CAPACITY.xlsx", col_types = "text")
)
                   
                   


```

## Summary

See what ranges of key variables we have in the tables

### Year


```{r Summary1}

for (aTable in names(myTables)){
  print(aTable)
  colToCheck <- 'YEAR'
  #print(colToCheck)
  print(ValuesPresentInData(myData=myTables[[aTable]], myColumn=colToCheck))
}


```

### Species

```{r Summary2}


for (aTable in names(myTables)){
  print(aTable)
  colToCheck <- 'SPECIES'
  #print(colToCheck)
  print(ValuesPresentInData(myData=myTables[[aTable]], myColumn=colToCheck))
}



```

## Table A Catch Summary

```{r TableA}

PlotTableA(DataToCheck = myTables[['Table_A']])


```

### Compare Table A to Tables C, D, E, and F and see if there is anything important missing

```{r TableA1}

kable(CompareTableA(DataToCheck = myTables[['Table_A']]), caption="Domain count comparison for top 10 species by value")

```

## Table C NAO OFR Discards age data


### Boxplot of ages

```{r TableCNAO1}

AgeBoxPlot(Data = myTables[['Table_C_NAO']], PlotTitle="Table C Ages")



```

### Numbers at age, by year and species (linear scale y axis)

```{r TableCNAO2}

#PlotTableC(LogYScale = FALSE)
PlotNoAtAge(LogYScale = FALSE,SpeciesToPlot = NULL, Data = myTables[['Table_C_NAO']], PlotTitle = "Table C Discards by age" )


```

### Numbers at age, by year and species (log scale y axis)

```{r TableCNAO3}

#PlotTableC(LogYScale = TRUE)
PlotNoAtAge(LogYScale = TRUE,SpeciesToPlot = NULL, Data = myTables[['Table_C_NAO']], PlotTitle = "Table C Discards by age" )

```

## Table D NAO OFR Discards length data

### Boxplot of lengths

```{r TableDNAO1}

LengthBoxPlot(Data = myTables[['Table_D_NAO']], PlotTitle="Table D Lengths")


```

### Numbers at length, by year and species (linear scale y axis)

```{r TableDNAO2}


# Too much data to have a single plot so we'll break it down into groups of species
mySpeciesList <- sort(unique(myTables[['Table_D_NAO']]$SPECIES))

numberInGroup <- 8.0

for(i in 1:ceiling(NROW(mySpeciesList)/numberInGroup)){
  
  firstValue <- 1 +   ((i-1)*numberInGroup) 
  secondValue<- firstValue +(numberInGroup -1)
  
  if (secondValue > NROW(mySpeciesList)){
    secondValue <- NROW(mySpeciesList)
  }
  
  #PlotTableD(LogYScale = FALSE, SpeciesToPlot = mySpeciesList[firstValue:secondValue])
  PlotNoAtLength(LogYScale = FALSE,SpeciesToPlot = mySpeciesList[firstValue:secondValue], Data = myTables[['Table_D_NAO']], PlotTitle = "Table D Discards by length" )
  
}



```


## Table E NAO OFR Landings age data

### Boxplot of Ages


```{r TableENAO1}

AgeBoxPlot(Data = myTables[['Table_E_NAO']], PlotTitle="Table E Ages")


```

### Numbers at age, by year and species (linear scale y axis)

```{r TableENAO2}


# Too much data to have a single plot so we'll break it down into groups of species
mySpeciesList <- sort(unique(myTables[['Table_E_NAO']]$SPECIES))

numberInGroup <- 8.0

for(i in 1:ceiling(NROW(mySpeciesList)/numberInGroup)){
  firstValue <- 1 +   ((i-1)*numberInGroup) 
  secondValue<- firstValue +(numberInGroup - 1)
  
  if (secondValue > NROW(mySpeciesList)){
    secondValue <- NROW(mySpeciesList)
  }
  #PlotTableE(LogYScale = FALSE, SpeciesToPlot = mySpeciesList[firstValue:secondValue])
  PlotNoAtAge(LogYScale = FALSE,SpeciesToPlot = mySpeciesList[firstValue:secondValue], Data = myTables[['Table_E_NAO']], PlotTitle = "Table E Landings by age" )
}


```

### Numbers at age, by year and species (log scale y axis)

```{r TableENAO3}


# Too much data to have a single plot so we'll break it down into groups of species
mySpeciesList <- sort(unique(myTables[['Table_E_NAO']]$SPECIES))

numberInGroup <- 8.0


for(i in 1:ceiling(NROW(mySpeciesList)/6.0)){
  firstValue <- 1 +   ((i-1)*6) 
  secondValue<- firstValue +5
  if (secondValue > NROW(mySpeciesList)){
    secondValue <- NROW(mySpeciesList)
  }
  #PlotTableE(LogYScale = TRUE, SpeciesToPlot = mySpeciesList[firstValue:secondValue])
  PlotNoAtAge(LogYScale = TRUE,SpeciesToPlot = mySpeciesList[firstValue:secondValue], Data = myTables[['Table_E_NAO']], PlotTitle = "Table E Landings by age" )
}



```

## Table F NAO OFR Landings length data

### Boxplot of lengths

```{r TableFNAO1}

LengthBoxPlot(Data = myTables[['Table_F_NAO']], PlotTitle="Table F Lengths")



```

### Numbers at length, by year and species (linear scale y axis)

```{r TableFNAO2}

# Too much data to have a single plot so we'll break it down into groups of species
mySpeciesList <- sort(unique(myTables[['Table_F_NAO']]$SPECIES))

numberInGroup <- 8.0

for(i in 1:ceiling(NROW(mySpeciesList)/numberInGroup)){
  firstValue <- 1 +   ((i-1)*numberInGroup) 
  secondValue<- firstValue + (numberInGroup - 1)
  
  if (secondValue > NROW(mySpeciesList)){
    secondValue <- NROW(mySpeciesList)
  }
  #PlotTableF(LogYScale = FALSE, SpeciesToPlot = mySpeciesList[firstValue:secondValue])
  PlotNoAtLength(LogYScale = FALSE,SpeciesToPlot = mySpeciesList[firstValue:secondValue], Data = myTables[['Table_F_NAO']], PlotTitle = "Table F Landings by length" )
}


```

### Numbers at length, by year and species (log scale y axis)

```{r TableFNAO3}


# Too much data to have a single plot so we'll break it down into groups of species
mySpeciesList <- sort(unique(myTables[['Table_F_NAO']]$SPECIES))

numberInGroup <- 8.0


for(i in 1:ceiling(NROW(mySpeciesList)/6.0)){
  firstValue <- 1 +   ((i-1)*6) 
  secondValue<- firstValue +5
  if (secondValue > NROW(mySpeciesList)){
    secondValue <- NROW(mySpeciesList)
  }
  #PlotTableF(LogYScale = TRUE, SpeciesToPlot = mySpeciesList[firstValue:secondValue])
  PlotNoAtLength(LogYScale = TRUE,SpeciesToPlot = mySpeciesList[firstValue:secondValue], Data = myTables[['Table_F_NAO']], PlotTitle = "Table F Landings by length" )
}

```

## Table G Effort Summary

```{r TableG1}

PlotTableG(DataToCheck = myTables[['Table_G']])

```

## Table H Landings by rectangle

Show the geographic range of landings by year

```{r TableH1}

plotTableH(DataToCheck = myTables[['Table_H']])

```

## Table I Effort by rectangle

Show the geographic range of effort by year

```{r TableI1}

plotTableI(DataToCheck = myTables[['Table_I']])

```



## Table J Capacity

```{r TableJ1}

plotTableJ(DataToCheck = myTables[['Table_J']])

```

